Step 1: Load and Explore the Dataset
Load the dataset into R.
```{r}
library(tidyr)
library(tidyverse)
superstore <- read.csv("C:/Users/Martijn/Downloads/drive-download-20250311T104144Z-001/Sample - Superstore/Sample - Superstore.csv")

```
Explore the dataset using str(), head(), and summary().
```{r}
print(str(superstore))
head(superstore)
summary(superstore)
```
Identify the number of rows and columns.
```{r}
ncol(superstore)
nrow(superstore)
```


Step 2: Basic Data Manipulation
Select the following columns: Order ID, Order Date, Customer Name, Sales, Profit.
```{r}
superstore_filter <- superstore %>%
  select(Order.ID, Order.Date, Customer.Name, Sales, Profit)
superstore_filter
```
Filter the dataset to show only orders with a profit greater than $100.
```{r}
superstore_filter <- superstore_filter %>%
  subset(superstore_filter$Profit > 100)
superstore_filter$Profit
```
Sort the dataset by Sales in descending order.
```{r}
superstore_filter <- superstore_filter %>%
  arrange(desc(Sales))
superstore_filter
```
Step 3: Handle Missing Data
Check for missing values in the dataset.
```{r}
sum(is.na(superstore))
```
Replace missing values in the Postal Code column with the mode (most frequent value).
```{r}
# Find the most frequent (mode) Postal.Code
mode_postal_code <- superstore %>%
  count(Postal.Code, sort = TRUE) %>%
  filter(!is.na(Postal.Code)) %>%
  slice(1) %>%
  pull(Postal.Code)

# Replace NA values with the most frequent Postal.Code
superstore <- superstore %>%
  mutate(Postal.Code = ifelse(is.na(Postal.Code), mode_postal_code, Postal.Code))
```
Remove rows with missing values in the Customer Name column.
```{r}
superstore <- superstore %>%
  subset(!is.na(Customer.Name))
```
Step 4: Create and Modify Columns
Create a new column Profit_Margin as the ratio of Profit to Sales.
```{r}
superstore <- superstore %>%
  mutate(Profit_Margin = Profit / Sales)
```
Create a new column Order_Year by extracting the year from Order Date.
```{r}
superstore <- superstore %>%
  mutate(Order_Year = str_extract(Order.Date,"\\d{4}"))
```
Convert the Order.Date column to a Date data type.
```{r}
superstore$Order.Date <- as.Date(superstore$Order.Date, format = "%m/%d/%Y")
str(superstore)
```
Step 5: Aggregating and Summarizing Data
Calculate the total sales and profit by Category.
```{r}
superstore_category <- superstore %>%
  group_by(Category) %>%
  summarize(Total.Sales =sum(Sales), Total.Profit = sum(Profit))
superstore_category
```
Find the average profit margin by Region.
```{r}
superstore %>%
  group_by(Region) %>%
  summarise(Average.Profit = mean(Profit_Margin))

```
Count the number of orders by Customer Segment.
```{r}
superstore %>%
  group_by(Segment) %>%
  summarise(NumberOfOrders = n())
```
Optional
Extra 1: Advanced Challenges
Identify and remove duplicate rows based on Order ID.
```{r}
#superstore[duplicated(superstore$Order.ID), ]
superstore <- superstore %>%
  distinct(Order.ID, .keep_all = TRUE)
```
Create a new column Discount_Level that categorizes discounts as "Low" (<10%), "Medium" (10-20%), and "High" (>20%).
```{r}
superstore <- superstore %>%
  mutate(Discount_Level = case_when(
    Discount < 0.1 ~ "Low",
    Discount > 0.1 & Discount < 0.2 ~ "Medium",
    Discount > 0.2 ~ "High"
    ))

```
Merge the dataset with a new dataset containing regional population data (create a dummy dataset for this purpose).
```{r}
df <- data.frame(
  Region = c("Central", "East", "South", "West"),
  Population = c(500000, 750000, 620000, 540000) # Example population numbers
)

superstore <- superstore %>%
  left_join(df, by = join_by(Region))

```
Extra 2: Data Visualization
Next class we will talk about data visualization, but let's see if you can pull it off on your own.
Create a bar plot of total sales by Category.
```{r}
f <- ggplot(superstore,aes(x=Category, y=Sales))
f 
+ geom_col(fill="skyblue", linetype=1) 
+ theme_classic() 
+ labs(title = "Sales by Category",
       x = "Product Category",
       y = "Total Sales")

```
Create a scatter plot of Sales vs. Profit with a trend line.
```{r}
s <- ggplot(superstore, aes(Sales, Profit))
s + geom_point(shape=1) + geom_smooth(color="blue", method="lm")
```
Create a histogram of Profit_Margin.
```{r}
h <- ggplot(superstore, aes(Profit_Margin))
h + geom_histogram(binwidth=0.1, fill="navyblue") + theme_classic() + labs(x = "Profit Margin", y="Count", title="Profit Margin Count")
```